LaMachine: A meta-distribution for NLP software

### 2019-01-28 docs/lamachine_clarin2018.tex
major edits in the description - attempt to move the technical details further behind

LaMachine is an open-source NLP software distribution. This implies the following in our case:

\begin{enumerate}
    \item LaMachine facilitates the installation, distribution and configuration of software. It does not fork, modify
        or appropriate the participating software in any way, nor does it provide a hosting place or repository for
        software.
    \item The software included in LaMachine must under a recognised (i.e. OSI-approved) open-source license.
        Proprietary (closed-source) software is explicitly excluded.
    \item The source
        code must be hosted in a public version controlled repository. \footnote{e.g. Github, Gitlab, Bitbucket, provided
        the respository is public}
    \item There must be a build process to compile the source, if applicable, and install the program or library.
    \item There must be some kind of release protocol (adhering to semantic versioning) that publishes software using the proper
technology-specific channels. We will elaborate on this later.
    \item All software that is incorporated in LaMachine must bear at least some relevance to the field of Natural Language Processing.
    \item LaMachine lives in an open-source ecosystem and therefore builds on Unix-like platforms; this primarily
        means Linux, as well as BSD and, with some restrictions, macOS. This by definition excludes certain software for different platforms, such as
        mobile platforms (Android/iOS/etc), native Windows/mac desktop applications, or certain interface types in
        general such as classical desktop GUI applications or mobile `apps', all of which fall beyond our scope.
        Cygwin\footnote{A unix environment on Windows} is not tested or
        supported either.  However, virtualisation technology enables deployment on a wider range of platforms, including Windows.
\end{enumerate}

We classify LaMachine as a \emph{meta distribution} as it can be installed in various contexts. The heart of LaMachine consists of a
set of machine-parsable instructions on how to obtain, build (e.g. compile from source), install and configure software.
These are implemented using Ansible\footnote{\url{https://www.ansible.com}; a popular open-source software provisioning,
configuration and deployment tool by Red Hat}.  This is notably different from the more
classical notion of Linux distributions, which generally provide their own repositories with (often binary) software
packages. LaMachine builds on this already established infrastructure by taking these repositories as a foundation where
possible. Similarly, as implied in point five above, there are different programming-language-specific ecosystems
providing their own repositories, such as the Python Package Index\footnote{\url{https://pypi.org}} for Python,
CRAN\footnote{\url{https://cran.r-project.org/}} for R, CPAN\footnote{\url{https://www.cpan.org}} for Perl, Maven
Central\footnote{\url{https://search.maven.org}} for Java.  LaMachine relies on those to pull and install software from
and never forks, archives, or modifies the software in any way. In doing so, we compel participating software projects
to adhere to well-established distribution standards and ensure the software is more sustainable towards the future
\cite{softwarequality}. Moreover, we ensure that LaMachine never becomes a prerequisite for the software but merely a
courtesy or convenience, and attempt to limit any amount of duplication in packaging and distribution efforts.

\section{Software}

@@

@@

(TODO) %TODO: some intro

For this workshop, computing power was purchased at SURFsara, a collaborative organization for ICT in Dutch education
and research. Subsequently, we had the ability to create Virtual Machines on their hosting platform. For this workshop
we had twelve participants and decided to create a single multi-core and high memory VM to share amongst the
participants.

@@

(TODO) %TODO: some intro

>>>>

\section{What is Lamachine}

LaMachine is an open-source NLP software distribution.  LaMachine facilitates the installation, distribution and configuration of software. It does not fork, modify or appropriate the participating software in any way, nor does it provide a hosting place or repository for software.
We classify LaMachine as a \emph{meta distribution} as it can be installed in various contexts. The heart of LaMachine consists of a
set of machine-parsable instructions on how to obtain, build (e.g. compile from source), install and configure software.
%iris:  eerst algemene uitleg over lamachine- dit is veel te gedetaileerd op dit punt-
%These are implemented using Ansible\footnote{\url{https://www.ansible.com}; a popular open-source software provisioning,
%configuration and deployment tool by Red Hat}.  
This is notably different from the more
classical notion of Linux distributions, which generally provide their own repositories with (often binary) software
packages. LaMachine builds on this already established infrastructure by taking these repositories as a foundation where
possible. Similarly, as implied in point five above, there are different programming-language-specific ecosystems
providing their own repositories, such as the Python Package Index\footnote{\url{https://pypi.org}} for Python,
CRAN\footnote{\url{https://cran.r-project.org/}} for R, CPAN\footnote{\url{https://www.cpan.org}} for Perl, Maven
Central\footnote{\url{https://search.maven.org}} for Java.  LaMachine relies on those to pull and install software from
and never forks, archives, or modifies the software in any way. In doing so, we compel participating software projects
to adhere to well-established distribution standards and ensure the software is more sustainable towards the future
\cite{softwarequality}. Moreover, we ensure that LaMachine never becomes a prerequisite for the software but merely a
courtesy or convenience, and attempt to limit any amount of duplication in packaging and distribution efforts.

LaMachine lives in an open-source ecosystem and therefore builds on Unix-like platforms; this primarily means Linux, as well as BSD and, with some restrictions, macOS. This by definition excludes certain software for different platforms, such as mobile platforms (Android/iOS/etc), native Windows/mac desktop applications, or certain interface types in general such as classical desktop GUI applications or mobile `apps', all of which fall beyond our scope. Cygwin\footnote{A unix environment on Windows} is not tested or
 supported either.  However, virtualisation technology enables deployment on a wider range of platforms, including Windows.

The software included in LaMachine has to live up (iris:ik zoek synoniem hier voor live up) to the following prerequisites:
\begin{enumerate}

    \item The software must have a recognised (i.e. OSI-approved) open-source license.
        Proprietary (closed-source) software is explicitly excluded.
    \item The source
        code must be hosted in a public version controlled repository. \footnote{e.g. Github, Gitlab, Bitbucket, provided
        the respository is public}
    \item There must be a build process to compile the source, if applicable, and install the program or library.
    \item There must be some kind of release protocol (adhering to semantic versioning) that publishes software using the proper
technology-specific channels. We will elaborate on this later.
    \item All software that is incorporated in LaMachine must bear at least some relevance to the field of Natural Language Processing.
    \item {\bf maarten help even met formuleren hier, ik bedoel dat zoiets als Pattern dus helaas niet mee kan doen omdat het niet meer onderhouden wordt }only maintained/uptodate date software that can compile using the current version of compilers and dependencies/libraries
\end{enumerate}

\section{ Included Software}

@@

In this section we present the technical design choices that were made.
%iris: Maarten je zou hier ook expliciet de motivatie voor de keuzen kunnen toevoegen

@@

The ICT-Research Platform Netherlands and NWO organize a yearly one-week workshop `ICT with Industry' \footnote{https://ict-research.nl/ict-with-industry/ictwi2019/} to stimulate collaboration between industry and academia. The industrical partner provides a problem and a team of researchers from different backgrounds and universities collaborate to come up with solutions. We partipated in the 2019 edition on the case study by the Dutch Royal Library who

For this workshop, computing power was purchased at SURFsara, a collaborative organization for ICT in Dutch education
and research. Subsequently, we had the ability to create Virtual Machines on their hosting platform. For this workshop
we had twelve participants and decided to create a single multi-core and high memory VM to share amongst the
participants to create one common digital workspace.

@@

Lamachine offered a convenient platform for a range of different explorations and experiments in the area of NLP and text mining. 
However, for some situations Lamachine was not a good fit for the tasks at hand: For team members who did not have experience with a non-Windows environment, Lamachine was not a suitable or useful tool. And also for members who wanted to use a graphical user interface for their text editors as we did not manage to get the x-forwarding working (note that this also depends on the overall VM). After a few attempts the team gave up on resolving this issue die to time pressure but it also demonstrated that finetuning the configuration of Lamachine on certain aspects is beyond the reach of a data scientist without system administration skills.
 %iris it moet nog wat beter geformuleerd worden maar je snaprt wat ik bedoel
